---
title: "Park Effect: Which Ballparks Helped the Most?"
subtitle: "Comparing home-run carry across MLB stadiums in 2025"
execute:
  echo: false
  warning: false
  message: false
---

::: {.lead}
The same swing does not travel the same way in every ballpark. A home run in one stadium can be a warning-track out in another, even when the exit velocity and launch angle look nearly identical.
:::

This page uses Statcast home-run data to compare how much a ballpark seems to boost or suppress home-run carry. Instead of treating every home run as identical, we compare each stadium’s median home-run distance to the season-wide median and ask: which parks were most generous to hitters in 2025?

```{r}
library(tidyverse)
library(plotly)

team_map <- c(
  "AZ" = "Arizona Diamondbacks", "ATH" = "Athletics", "ATL" = "Atlanta Braves",
  "BAL" = "Baltimore Orioles", "BOS" = "Boston Red Sox", "CHC" = "Chicago Cubs",
  "CWS" = "Chicago White Sox", "CIN" = "Cincinnati Reds", "CLE" = "Cleveland Guardians",
  "COL" = "Colorado Rockies", "DET" = "Detroit Tigers", "HOU" = "Houston Astros",
  "KC" = "Kansas City Royals", "LAA" = "Los Angeles Angels", "LAD" = "Los Angeles Dodgers",
  "MIA" = "Miami Marlins", "MIL" = "Milwaukee Brewers", "MIN" = "Minnesota Twins",
  "NYM" = "New York Mets", "NYY" = "New York Yankees", "PHI" = "Philadelphia Phillies",
  "PIT" = "Pittsburgh Pirates", "SD" = "San Diego Padres", "SF" = "San Francisco Giants",
  "SEA" = "Seattle Mariners", "STL" = "St. Louis Cardinals", "TB" = "Tampa Bay Rays",
  "TEX" = "Texas Rangers", "TOR" = "Toronto Blue Jays", "WSH" = "Washington Nationals"
)

home_runs <- read_csv("data/home_runs_2025.csv", show_col_types = FALSE)
ballparks <- read_csv("data/ballpark_dimensions.csv", show_col_types = FALSE)
league_median <- median(home_runs$hit_distance_sc, na.rm = TRUE)

park_summary <- home_runs |>
  filter(!is.na(hit_distance_sc)) |>
  mutate(team = unname(team_map[home_team])) |>
  filter(!is.na(team)) |>
  group_by(team) |>
  summarise(
    home_runs = n(),
    median_distance = median(hit_distance_sc),
    adjustment = median_distance - league_median,
    .groups = "drop"
  ) |>
  left_join(ballparks |> select(team, venue), by = "team") |>
  arrange(median_distance)

p <- ggplot(park_summary, aes(
  x = median_distance,
  y = reorder(venue, median_distance),
  fill = adjustment,
  text = paste0(
    "<b>", venue, "</b><br>",
    "Median HR Distance: ", round(median_distance), " ft<br>",
    "Park Adjustment: ", if_else(adjustment >= 0, "+", ""), round(adjustment), " ft"
  )
)) +
  geom_col(width = 0.8) +
  scale_fill_gradient2(low = "#2c7fb8", mid = "#f7f7f7", high = "#d95f4b", midpoint = 0) +
  labs(
    title = "Which parks were most generous to hitters?",
    x = "Median home-run distance (feet)",
    y = NULL,
    fill = "Adjustment (ft)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

p |>
  ggplotly(tooltip = "text") |>
  layout(autosize = TRUE) |>
  config(responsive = TRUE, scrollZoom = FALSE)
```

## What this plot is showing

This chart compares each stadium’s median home-run distance against the overall 2025 median. Positive values indicate a park where home runs tended to carry farther than the league average; negative values suggest a tougher environment for hitters.

## The most hitter-friendly ballparks

These are the stadiums where the median home run traveled furthest above the league norm.

```{r}
#| results: asis

top_parks <- park_summary |>
  arrange(desc(median_distance)) |>
  slice_head(n = 8) |>
  transmute(
    Ballpark = venue,
    `Home runs` = home_runs,
    `Median distance` = round(median_distance),
    `Adjustment vs league median` = round(adjustment)
  )

knitr::kable(top_parks, format = "html")
```
